---
title: "Esercizi_capitolo5"
author: "Alessandro Dell'Orto"
date: "12/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Control flow

\stepcounter{section}
## Choices


__[Q1]{.Q}__: What type of vector does each of the following calls to `ifelse()` return?

```{r, eval = FALSE}
ifelse(TRUE, 1, "no")
ifelse(FALSE, 1, "no")
ifelse(NA, 1, "no")
```

__[A]{.solved}__: 

```{r}
?ifelse  ##ifelse(test, yes, no) returns a value with the same shape as test which is filled with elements selected from either yes or no depending on whether the element of test is TRUE or FALSE.
##       | yes if x=TRUE
## f(x)= | no  if x=FALSE NB typeof(x) is coerced to logical
##       | NA  if x=NA

##without coercion
ifelse(TRUE, 1, "no")
ifelse(FALSE, 1, "no")
ifelse(NA, 1, "no")
##with coercion
ifelse(0,'yes','no')
ifelse(1,'yes','no')
ifelse('lept1','yes','no')
ifelse(2.5,'yes','no')
ifelse(-2.5,'yes','no')
ifelse(-2L,'yes','no')
ifelse(2L,'yes','no')
ifelse(0L,'yes','no')
ifelse(0.5,'yes','no')
ifelse(0.0,'yes','no')
ifelse(c(),'yes','no')
ifelse(c(0),'yes','no')
ifelse(c(0,1,3),'yes','no')
ifelse(list(0,1,3),'yes','no')

```

__[Q2]{.Q}__: Why does the following code work?

```{r}
x <- 1:10
if (length(x)) "not empty" else "empty"

x <- numeric()
if (length(x)) "not empty" else "empty"
```

__[A]{.solved}__: lenght(x) is coerced to logical
```{r}
x <- 1:10
if (length(x)) "not empty" else "empty"
(as.logical(length(x)))

x <- 0
if (length(x)) "not empty" else "empty"
(as.logical(length(x)))

x <- numeric()
if (length(x)) "not empty" else "empty"
(as.logical(length(x)))
```
## Loops

__[Q1]{.Q}__: Why does this code succeed without errors or warnings?

```{r, results = FALSE}
x <- numeric()
out <- vector("list", length(x))
for (i in 1:length(x)) {
  out[i] <- x[i] ^ 2
}
out
```

__[A]{.solved}__: print it

```{r, results = FALSE}
x <- numeric() 
cat('lunghezza x = ',length(x), "\n")
print(x)
out <- vector("list", length(x))
cat('lunghezza out = ',length(out), "\n")
#tracemem(out)
for (i in 1:length(x)) {   ##from 1 to 0
  cat('i = ',i, "\n")
  cat('x[i] = ',x[i], "\n")
  out[i] <- x[i] ^ 2       ## x[1] is NA (it does not exist), x[0] is blank and add an empty element to a list does not affect its length
}
print(out)
```

__[Q2]{.Q}__: When the following code is evaluated, what can you say about the vector being iterated?

```{r}
xs <- c(1, 2, 3)
for (x in xs) {
  xs <- c(xs, x * 2)
}
xs
```

__[A]{.solved}__: print it

```{r}
xs <- c(1, 2, 3)
tracemem(xs)
for (x in xs) {     ## x takes the values from xs before the loop and stores them
  print(x)
  xs <- c(xs, x * 2)
}
xs
```

__[Q3]{.Q}__: What does the following code tell you about when the index is updated?

```{r}
for (i in 1:3) {
  i <- i * 2
  print(i) 
}
```

__[A]{.solved}__: As before. i takes the values before the loop and stores them and they can't be modified