---
title: "Esercizi_capitolo3"
author: "Alessandro Dell'Orto"
date: "12/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Subsetting 

## Selecting multiple elements

__[Q1]{.Q}__: Fix each of the following common data frame subsetting errors:

```{r, eval = FALSE}
mtcars[mtcars$cyl = 4, ]

mtcars[-1:4, ]

mtcars[mtcars$cyl <= 5]

mtcars[mtcars$cyl == 4 | 6, ]

```
__[A]{.solved}__: 
```{r, eval = FALSE}
mtcars[mtcars$cyl == 4, ] ## ==

mtcars[1:4, ] ## or all negative or all positive

mtcars[mtcars$cyl <= 5,] ## missing ,

mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ] ##second condition

```



__[Q2]{.Q}__: Why does the following code yield five missing values? (Hint: why is it different from `x[NA_real_]`?)

```{r}
x <- 1:5
x[NA]
```

__[A]{.solved}__: NA is logical, and with x[NA] we are creating a subset using a logical vector

```{r}
typeof(NA)

x <- 1:5

x[NA]
x[TRUE]
x[FALSE]

```

__[Q3]{.Q}__: What does `upper.tri()` return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?

```{r echo = TRUE, results = 'hide'}
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]
```

__[A]{.solved}__: let's see
```{r}
?upper.tri ## cit. Returns a matrix of logicals the same size of a given matrix with entries TRUE in the lower or upper triangle.
x
upper.tri(x)

##i don't get the last question
```


__[Q4]{.Q}__: Why does `mtcars[1:20]` return an error? How does it differ from the similar `mtcars[1:20, ]`?
__[A]{.solved}__: with df[] you are selecting columns, while df[,] you are selecting rows 
```{r}
mtcars ## this is the df. only 11 columns
```
```{r}
mtcars[1:2] ##select only col 1 and 2
mtcars[1:2,]#select only row 1 and 2
```



__[Q5]{.Q}__: Implement your own function that extracts the diagonal entries from a matrix (it should behave like `diag(x)` where `x` is a matrix).
__[A]{.solved}__: let's try
```{r}
diagonalley <- function(M) {
  n<-min(dim(M))
  delta<-1:n
  M[cbind(delta,delta)]
}
# Let's check if it works
MM <- matrix(1:48, 6)
print(MM)
diag(MM)
diagonalley(MM)
```


__[Q6]{.Q}__: What does `df[is.na(df)] <- 0` do? How does it work?

__[A]{.solved}__: imputation! nan become 0. first use a logical vector to subset only nan then assign them 0 value

## Selecting a single element

__[Q1]{.Q}__: Brainstorm as many ways as possible to extract the third value from the `cyl` variable in the `mtcars` dataset.

__[A]{.solved}__:

```{r}

mtcars$cyl[3]
mtcars$cyl[[3]]
mtcars[3,"cyl"]
mtcars[[3,"cyl"]]
mtcars[3,][["cyl"]]
mtcars[3,]["cyl"][[1]]
```

__[Q2]{.Q}__: Given a linear model, e.g. `mod <- lm(mpg ~ wt, data = mtcars)`, extract the residual degrees of freedom. Extract the R squared from the model summary (`summary(mod)`).

__[A]{.solved}__: 

```{r}
?lm  ## An object of class "lm" is a list containing at least the following components:...
mod <- lm(mpg ~ wt, data = mtcars)

head(mod)
```

```{r}
mod$df.residual
summary(mod)$r.squared
```

\stepcounter{section}
## Applications

__[Q1]{.Q}__: How would you randomly permute the columns of a data frame? (This is an important technique in random forests.) Can you simultaneously permute the rows and columns in one step?

__[A]{.solved}__: thanks to stackoverflow:

```{r,eval = FALSE}
mtcars[sample(nrow(mtcars)),] #permute rows
mtcars[sample(ncol(mtcars))]  #permute col
mtcars[sample(nrow(mtcars)),sample(ncol(mtcars))] #permute row and col

```

__[Q2]{.Q}__: How would you select a random sample of `m` rows from a data frame? What if the sample had to be contiguous (i.e. with an initial row, a final row, and every row in between)?

__[A]{.solved}__: first question 

```{r, eval = FALSE}
?sample ##sample(x, size,...)
m <- 5
mtcars[sample(nrow(mtcars), m), ]
```
second question

```{r, eval = FALSE}
randrow <- as.integer(runif(1,min=1,max=nrow(mtcars)-m))  ## non mi piace
mtcars[randrow:(randrow+m-1),]
```

__[Q3]{.Q}__: How could you put the columns in a data frame in alphabetical order?

__[A]{.solved}__:
```{r, eval = FALSE}
head(mtcars)
nomi<-names(mtcars)
ordnom<-order(nomi)
head(mtcars[ordnom])
```