---
title: "Studio di R"
author: "Alessandro Dell'Orto"
date: "November 30, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 2

Given the following data frame, how do I create a new column called “3” 
that contains the sum of 1 and 2? You may only use $, not [[. 
What makes 1, 2, and 3 challenging as variable names?
```{r}
df <- data.frame(runif(3), runif(3))
names(df) <- c(1, 2)
#### my answer
df[3]<-df[1]+df[2]
names(df[3])<-3
print(df)
### MARIO Puoi provare a rifarlo usando $ come chiesto nell'esempio
### attenzione ad usare quel formalismo che può essere fuorviante
#guarda questo codice
df <- data.frame(runif(3), runif(3))
names(df) <- c(2, 1)
df[3]<- df[1] - df[2]
names(df[3])<-3
print(df)
# dobbiamo essere sicuri non solo che il codice funzioni ma che risponda ai cambiamenti che facciamo nel modo giusto
```



In the following code, how much memory does y occupy?
```{r}
x <- runif(1e6)
y <- list(x, x, x)
### #my answer
object.size(y)
```



On which line does a get copied in the following example?
```{r}
a <- c(1, 5, 3, 2)
b <- a
b[[1]] <- 10
#### my answer
#### at the last. i used lobstr::obj_addr to check
```



Explain the relationship between a, b, c and d in the following code:
```{r}
a <- 1:10
b <- a
c <- b
d <- 1:10
####my answer
#### a,b,c point to the same address. 
lobstr::obj_addr(a)==lobstr::obj_addr(b) & lobstr::obj_addr(a)==lobstr::obj_addr(c)
lobstr::obj_addr(d)!=lobstr::obj_addr(a)
```



The following code accesses the mean function in multiple ways. 
Do they all point to the same underlying function object? Verify this with lobstr::obj_addr().
```{r}
mean
base::mean
get("mean")
evalq(mean)
match.fun("mean")
####my answer
lobstr::obj_addr(mean)==lobstr::obj_addr(base::mean) & lobstr::obj_addr(mean)==lobstr::obj_addr(get("mean")) &
lobstr::obj_addr(mean)==lobstr::obj_addr(evalq(mean)) & lobstr::obj_addr(mean)==lobstr::obj_addr(match.fun("mean")) 
```



By default, base R data import functions, like read.csv(), will automatically convert non-syntactic names 
to syntactic ones. Why might this be problematic? What option allows you to suppress this behaviour?
```{r}
#### my answer. cit [stackoverflow]
?check.names
#### check.names: logical.  If ‘TRUE’ then the names of the variables in the
#### data frame are checked to ensure that they are syntactically valid variable names. 
```




What rules does make.names() use to convert non-syntactic names into syntactic ones?
```{r}
#### my answer. cit [stackoverflow]
#### The character "X" is prepended if necessary. All invalid characters are translated to ".". 
#### A missing value is translated to "NA". Names which match R keywords have a dot appended to them.
#### Duplicated values are altered by make.unique
?make.names
```



I slightly simplified the rules that govern syntactic names. Why is .123e1 not a syntactic name? 
Read ?make.names for the full details.
```{r}
#### my answer
#### it begins with ""." . let's try.
make.names(.123e1)
```



Why is tracemem(1:10) not useful?
Because 1:10 is on a fixed memory address (and this address cannot change), i.e. 1:10 is not a variable



Explain why tracemem() shows two copies when you run this code. Hint: carefully look at the difference between this code and the code shown earlier in the section.
```{r}
x <- c(1L, 2L, 3L)
tracemem(x)

x[[3]] <- 4

### my answer 
### before x is an integer then a double

```



What happens when you run this code?
```{r}
x <- list(1:10)
x[[2]] <- x
###my answer
### we can simply use ref()
ref(x)
### MARIO ref() viene da un pacchetto? Non la conosco come funzione base di R e non mi risulta. Poichè viene da un pacchetto che non mi sembra carichi
### all'inizio del RMarkdown o library(lobstr) o lobstr::ref(). Funziona questo markdown?
```



In the following example, why are object.size(y) and obj_size(y) so radically different? Consult the documentation of object.size().
```{r}
y <- rep(list(runif(1e4)), 100)

object.size(y)
#> 8005648 bytes
obj_size(y)
#> 80,896 B
###my answer
### obj_size counts shared item only one time
```

Take the following list. Why is its size somewhat misleading?
```{r}
funs <- list(mean, sd, var)
obj_size(funs)
#> 17,608 B
### my answer
### i read the solution.
## MARIO e che c'era scritto?
```

Predict the output of the following code:
```{r}
a <- runif(1e6)
lobstr::obj_size(a)

b <- list(a, a)
lobstr::obj_size(b)
lobstr::obj_size(a, b)

b[[1]][[1]] <- 10
lobstr::obj_size(b)
lobstr::obj_size(a, b)

b[[2]][[1]] <- 10
lobstr::obj_size(b)
lobstr::obj_size(a, b)
```


Explain why the following code doesn’t create a circular list.
```{r}
x <- list()
x[[1]] <- x

###my answer
### because R has the copy on modify.
```

Wrap the two methods for subtracting medians into two functions, then use the ‘bench’ package17 to carefully compare their speeds. How does performance change as the number of columns increase?
I dont know how


What happens if you attempt to use tracemem() on an environment?
It won't work on env.
MARIO Why?
